---
title: "Regression Project"
output: pdf_document
date: "2023-10-15"
---

```{r warning=FALSE}
library(readr)
library(GWmodel)
library(nortest)
library(lmtest)
library(car)
```

# kemiskinan2.xlsx

```{r}
library(readxl)
data <- read_excel("C:/Users/evopa/Downloads/kemiskinan2.xlsx")
data
```



`


```{r}
par(mfrow = c(4, 2))

boxplot(data$education_completion_rate, main="Education Completion Rate", col="lightblue", border="black")
boxplot(data$adequate_sanitation, main="Adequate Sanitation", col="lightgreen", border="black")
boxplot(data$drinking_water, main="Drinking Water", col="lightcoral", border="black")
boxplot(data$health_facilities, main="Health Facilities", col="lightyellow", border="black")
boxplot(data$own_house, main="Own House", col="lightsalmon", border="black")
boxplot(data$electricity_pln, main="Electricity PLN", col="lightcyan", border="black")
boxplot(data$poverty, main="Poverty", col="lightblue", border="black")


```
par(mfrow = c(2, 4))

boxplot(data$education_completion_rate, main="Education Completion Rate", col="lightblue", border="black")
boxplot(data$adequate_sanitation, main="Adequate Sanitation", col="lightgreen", border="black")
boxplot(data$drinking_water, main="Drinking Water", col="lightcoral", border="black")
boxplot(data$health_facilities, main="Health Facilities", col="lightyellow", border="black")
boxplot(data$own_house, main="Own House", col="lightsalmon", border="black")
boxplot(data$electricity_pln, main="Electricity PLN", col="lightcyan", border="black")
boxplot(data$poverty, main="Poverty", col="lightblue", border="black")

## Linear Regression OLS

```{r}
modelOLS <- lm(formula = poverty ~ education_completion_rate + adequate_sanitation + drinking_water + health_facilities + own_house + electricity_pln, data = data)
sum = summary(modelOLS)
sum
```

```{r}
cat("AIC = ",AIC(modelOLS))
cat("\nR2 = ",sum$r.squared)
```

### Pearson Correlation Test

```{r}
cor.test(data$poverty, data$drinking_water)
```

### Model Adequacy Checking

#### Normality Test

Hypothesis:

$$
\begin{aligned}
H_0&: \text{Errors are distributed normally}\\
H_1&: \text{Errors are not distributed normally}
\end{aligned}
$$

Significance Level:

$$
\alpha = 5\%=0.05
$$

P-value vs $\alpha$:

```{r}
error <- resid(modelOLS)
error

lillie.test(error)
```

Since the p-value \> 0.05, we fail to reject $H_0$.

Thus, the errors are distributed normally

#### Homoscedasticity Test

Hypothesis:

$$
\begin{aligned}
H_0&: \text{The residuals are distributed with equal variance (Homoscedastic)}\\
H_1&: \text{The residuals are distributed with non equal variance (Heteroscedastic)}
\end{aligned}
$$

Significance Level:

$$
\alpha = 5\%=0.05
$$

P-value vs $\alpha$:

```{r}
bptest(modelOLS)
```

Since the p-value \< 0.05, we reject $H_0$.

Thus, the residuals are distributed with non equal variance. Jadi tidak lolos uji ini.

#### Autocorrelation Test

Hypothesis:

$$
\begin{aligned}
H_0&: \text{The residuals are non autocorrelated}\\
H_1&: \text{The residuals are autocorrelated}
\end{aligned}
$$

Significance Level:

$$
\alpha = 5\%=0.05
$$

P-value vs $\alpha$:

```{r}
dwtest(modelOLS)
```

Since the p-value \> 0.05, we fail to reject $H_0$.

Thus, the errors are non autocorrelated

### Multicollinearity Check

```{r}
vif(modelOLS)
```

Since there is no value of VIF that greater than 10, we can conclude that there are no multicollinearity between the variables.

### AIC, BIC, AICc, and SSE
```{r}
AIC(modelOLS)
BIC(modelOLS)
AICc(modelOLS)
```
```{r}
anova(modelOLS)
```
```{r}
# high leverage => 2(p+1)/n = 2(7)/34 = 0.4118
hats <- as.data.frame(gwr.fit$lhat)
hats%>%arrange(desc(hats))
str(gwr.fit)

#outlier in y
boxplot(data$Y)
hist(data$Y)

#influential obs -> ci > 1
c <- as.data.frame(cooks.distance(model))
c%>%arrange(desc(c))
```
## Variable Selection: Forward
```{r}
mfor <- step(lm(poverty~1, data=data), direction="forward", scope=~education_completion_rate + adequate_sanitation + drinking_water + health_facilities + own_house + electricity_pln)
```
```{r}
model2=lm(poverty ~ drinking_water + adequate_sanitation + education_completion_rate, data=data)
summary(model2)
```
```{r}
AIC(model2)
BIC(model2)
AICc(model2)
```
```{r}
anova(model2)
```

## Variable Selection: Backward
```{r}
mback <- step(lm(poverty~education_completion_rate + adequate_sanitation + drinking_water + health_facilities + own_house + electricity_pln, data=data), direction="backward")
```

## Weighted Least Square Regression

Since the OLS model does not satisfy all of the model adequacy check, we try to use the weighted least square regression to fix the homoscedastic issue.

```{r}
wt <- 1/lm(abs(modelOLS$residuals) ~ modelOLS$fitted.values)$fitted.values^2
wls_model <- lm(formula = poverty ~ education_completion_rate + adequate_sanitation + drinking_water + health_facilities + own_house + electricity_pln, data = data, weights=wt)
summary(wls_model)
```
```{r}
AIC(wls_model)
BIC(wls_model)
AICc(wls_model)
```
```{r}
anova(wls_model)
```

Since the $R^2$ is below the OLS model (as benchmark), we try to use another method

## Geographically Weighted Regression

```{r warning=FALSE}
library(spgwr)
library(spdep)
library(spatialreg)
library(gwrr)
library(sf)
library(mapview)
library(tibble)
```

```{r}
df <- as_tibble(data)
df_spasial <- st_as_sf(df, coords = c("longitude", "latitude"), crs = 4326)
df_spasial_sp <- as(df_spasial, "Spatial")
df_spasial_sp
```

```{r}
#coords <- coordinates(df_spasial_sp)
#bobot <- nb2listw(knn2nb(knearneigh(coords)))
#moran.test(df_spasial_sp$y, ww, alternative="greater")
#euclidean <- gw.dist(coords)
#View(euclidean)
```

### Finding the Optimal Bandwidth

```{r}
gwr_bandwidth <- gwr.sel(poverty ~ education_completion_rate + adequate_sanitation + drinking_water + health_facilities + 
                           own_house + electricity_pln, df_spasial_sp, gweight = gwr.Gauss)
gwr_bandwidth
```

### The GWR Model

```{r}
gwr.fit <- gwr(poverty ~ education_completion_rate + adequate_sanitation + drinking_water + health_facilities + own_house + 
                 electricity_pln, data = df_spasial_sp, bandwidth = gwr_bandwidth, hatmatrix=TRUE, se.fit=TRUE, gweight = gwr.Gauss)
gwr.fit
```
```{r}
BIC(gwr.fit)
```
```{r}
gwr_bandwidth2 <- gwr.sel(poverty ~ education_completion_rate + adequate_sanitation + drinking_water, df_spasial_sp, gweight = gwr.Gauss)
gwr_bandwidth2
```
```{r}
gwrfit_reduced=gwr(poverty ~ education_completion_rate + adequate_sanitation + drinking_water, data = df_spasial_sp, bandwidth = gwr_bandwidth2, hatmatrix=TRUE, se.fit=TRUE, gweight = gwr.Gauss)
gwrfit_reduced
```

```{r}
model = c("OLS","GWR")
R2 = c(sum$r.squared, 0.8194663)
R2
```

```{r}
AIC = c(AIC(modelOLS),gwr.fit$results$AICh)
AIC
```

```{r}
evaluasi = data.frame(model,R2,AIC)
evaluasi
```

# kemiskinan.xlsx

```{r}
library(readxl)
data <- read_excel("C:/Users/evopa/Downloads/kemiskinan.xlsx")
data
```

## Linear Regression OLS

```{r}
modelOLS <- lm(formula = miskin ~ tingkat_penyelesaian_pendidikan + sanitasi_layak + luas_lantai + keluhan_kesehatan + air_minum + fasilitas_kesehatan +rumah_sendiri, data = data)
sum = summary(modelOLS)
sum
```

```{r}
cat("AIC = ",AIC(modelOLS))
cat("\nR2 = ",sum$r.squared)
```

### Pearson Correlation Test

```{r}
cor.test(data$miskin, data$air_minum)
```

### Model Adequacy Checking

#### Normality Test

Hypothesis:

$$
\begin{aligned}
H_0&: \text{Errors are distributed normally}\\
H_1&: \text{Errors are not distributed normally}
\end{aligned}
$$

Significance Level:

$$
\alpha = 5\%=0.05
$$

P-value vs $\alpha$:

```{r}
error <- resid(modelOLS)
error

lillie.test(error)
```

Since the p-value \> 0.05, we fail to reject $H_0$.

Thus, the errors are distributed normally

#### Homoscedasticity Test

Hypothesis:

$$
\begin{aligned}
H_0&: \text{The residuals are distributed with equal variance (Homoscedastic)}\\
H_1&: \text{The residuals are distributed with non equal variance (Heteroscedastic)}
\end{aligned}
$$

Significance Level:

$$
\alpha = 5\%=0.05
$$

P-value vs $\alpha$:

```{r}
bptest(modelOLS)
```

Since the p-value \< 0.05, we reject $H_0$.

Thus, the residuals are distributed with non equal variance. Jadi tidak lolos uji ini.

#### Autocorrelation Test

Hypothesis:

$$
\begin{aligned}
H_0&: \text{The residuals are non autocorrelated}\\
H_1&: \text{The residuals are autocorrelated}
\end{aligned}
$$

Significance Level:

$$
\alpha = 5\%=0.05
$$

P-value vs $\alpha$:

```{r}
dwtest(modelOLS)
```

Since the p-value \> 0.05, we fail to reject $H_0$.

Thus, the errors are non autocorrelated

### Multicollinearity Check

```{r}
vif(modelOLS)
```

Since there is no value of VIF that greater than 10, we can conclude that there are no multicollinearity between the variables.

## Weighted Least Square Regression

Since the OLS model does not satisfy all of the model adequacy check, we try to use the weighted least square regression to fix the homoscedastic issue.

```{r}
wt <- 1/lm(abs(modelOLS$residuals) ~ modelOLS$fitted.values)$fitted.values^2
wls_model <- lm(formula = miskin ~ tingkat_penyelesaian_pendidikan + sanitasi_layak + luas_lantai + keluhan_kesehatan + air_minum + fasilitas_kesehatan +rumah_sendiri, data = data, weights=wt)

summary(wls_model)
```

Since the $R^2$ is below the OLS model (as benchmark), we try to use another method

## Geographically Weighted Regression

```{r warning=FALSE}
library(spgwr)
library(spdep)
library(spatialreg)
library(gwrr)
library(sf)
library(mapview)
library(tibble)
```

```{r}
df <- as_tibble(data)
df_spasial <- st_as_sf(df, coords = c("longitude", "latitude"), crs = 4326)
df_spasial_sp <- as(df_spasial, "Spatial")
df_spasial_sp
```

```{r}
#coords <- coordinates(df_spasial_sp)
#bobot <- nb2listw(knn2nb(knearneigh(coords)))
#moran.test(df_spasial_sp$y, ww, alternative="greater")
#euclidean <- gw.dist(coords)
#View(euclidean)
```

### Finding the Optimal Bandwidth

```{r}
gwr_bandwidth <- gwr.sel(miskin ~ tingkat_penyelesaian_pendidikan + sanitasi_layak + luas_lantai + keluhan_kesehatan + air_minum + fasilitas_kesehatan +rumah_sendiri, df_spasial_sp, gweight = gwr.Gauss)
gwr_bandwidth
```

### The GWR Model

```{r}
gwr.fit <- gwr(miskin ~ tingkat_penyelesaian_pendidikan + sanitasi_layak + luas_lantai + keluhan_kesehatan + air_minum + fasilitas_kesehatan +rumah_sendiri, data = df_spasial_sp, bandwidth = gwr_bandwidth, hatmatrix=TRUE, se.fit=TRUE, gweight = gwr.Gauss)
gwr.fit
```

```{r}
model = c("OLS","GWR")
R2 = c(sum$r.squared, 0.8330831)
R2
```

```{r}
AIC = c(AIC(modelOLS),gwr.fit$results$AICh)
AIC
```

```{r}
evaluasi = data.frame(model,R2,AIC)
evaluasi
```
